---
title: "Homework 5"
author: "Shritama Ray"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output: github_document

---

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)
library(broom)
library(p8105.datasets)
library(viridis)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

### Problem 1

The code chunk below imports the data in individual spreadsheets contained in `./data/zip_data/`. To do this, I create a dataframe that includes the list of all files in that directory and the complete path to each file. As a next step, I `map` over paths and import data using the `read_csv` function. Finally, I `unnest` the result of `map`.

```{r}
full_df = 
  tibble(
    files = list.files("Data/zip_data/"),
    path = str_c("Data/zip_data/", files)
  ) %>% 
  mutate(data = map(path, read_csv)) %>% 
  unnest()
```

The result of the previous code chunk isn't tidy -- data are wide rather than long, and some important variables are included as parts of others. The code chunk below tides the data using string manipulations on the file, converting from wide to long, and selecting relevant variables. 

```{r}
tidy_df = 
  full_df %>% 
  mutate(
    files = str_replace(files, ".csv", ""),
    group = str_sub(files, 1, 3)) %>% 
  pivot_longer(
    week_1:week_8,
    names_to = "week",
    values_to = "outcome",
    names_prefix = "week_") %>% 
  mutate(week = as.numeric(week)) %>% 
  select(group, subj = files, week, outcome)
```

Finally, the code chunk below creates a plot showing individual data, faceted by group. 

```{r}
tidy_df %>% 
  ggplot(aes(x = week, y = outcome, group = subj, color = group)) + 
  geom_point() + 
  geom_path() + 
  facet_grid(~group)
```

This plot suggests high within-subject correlation -- subjects who start above average end up above average, and those that start below average end up below average. Subjects in the control group generally don't change over time, but those in the experiment group increase their outcome in a roughly linear way.  

### Problem 2: Washington Post Homicides Data

**Read in & tidy the data:**
```{r read and tidy data}
#Read in data from github
homicides = read.csv('https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv') %>%
  janitor::clean_names()%>%
  mutate(
    city_state = str_c(city, ", ", state)
  )
```
The raw data set has *`r nrow(homicides)`* observations of *`r ncol(homicides)`* variables. It contains data on homicides, with variables such as  reported date,  victim's name,  race/age/sex,  disposition, city and state.

**Summary of Unsolved Homicides:**
```{r city summary}
homicides %>%
  filter(disposition == "Closed without arrest" | disposition == "Open/No arrest")%>%
  group_by(city_state, disposition)%>%
  count() %>%
  knitr::kable()
```

**Baltimore, MD Proportions**
Use the prop.test function to estimate the proportion of homicides that are unsolved; save the output of prop.test as an R object, apply the broom::tidy to this object and pull the estimated proportion and confidence intervals from the resulting tidy dataframe

```{r baltimore}
baltimore =
  homicides %>%
  filter(city_state == "Baltimore, MD") %>%
  mutate(solved = case_when(
           disposition == "Closed without arrest" ~ "no",
           disposition == "Open/No arrest" ~ "no",
           disposition == "Closed by arrest" ~ "yes")) %>%
  summarize(
    unsolved = sum(solved == "no"),
    total_balt = n()) 

#perform test & save in dataframe
balt_prop_test = prop.test(baltimore$unsolved, baltimore$total_balt) 

# pull estimated proportion & confidence interval
balt_prop_test %>%
  broom::tidy()%>%
  select(estimate, conf.low, conf.high)

#pull() only works for 1 variable at a time 
```

**prop.test for all cities**

```{r prop test function}

#create a function to perform prop.test()

fxn_prop_test = function(city_data)
```

### Problem 3: Power in a 1-sample t-test

n = 30
sigma = 5
mu = 0

**Generate 5000 datasets from the model: x ~ Normal[mu, sigma]:**
```{r sim mean sd}

#create function
sim_mean_pval = function(n, mu, sigma) {
  sim_data = tibble(
    x = rnorm(n, mean = mu, sd = sigma),
  )
  
  sim_data %>% 
    summarize(
      mu_hat = mean(x),
      t.test(x, mu = mu, alternative = "less", conf.level = 0.95) %>%
        broom::tidy() %>%
        select(p.value)
    )
}

#run function 5000 times and save results to a list
output = vector("list", 100)

for (i in 1:100) {
  output[[i]] = sim_mean_pval(n = 30, mu = 0, sigma = 5)
}

sim_results = bind_rows(output)
  
```

**Repeat the above for mu = 1, 2, 3, 4, 5, 6**
```{r varying mu}

#create a dataframe with all the outputs
t_test_stats =
tibble(
  mu = c(1:6)
) %>%
  mutate(
    stats = map(.x = mu, ~rerun(100, sim_mean_pval(n = 30, mu = .x, sigma = 5)))
  ) %>%
unnest(stats) %>%
unnest(stats)

```

**Plot proportion of times null was rejected vs. mu**
```{r plot power}
t_test_stats %>%
  mutate(
  reject = (p.value < 0.05)) %>%
  group_by(mu) %>%
    summarize(
    power = sum(reject)/length(p.value)) %>%
  ggplot(aes(x = mu, y = power)) + geom_line() +
  xlab("True Mean")+
  ylab("Test Power")
```

**Plot the average estimate of μ-hat vs. the true value of μ. Make a second plot of the average estimate of μ-hat only in samples for which the null was rejected vs. the true value of μ.**
```{r plot mu hat}
plot_mu_hat = 
  t_test_stats %>%
  group_by(mu) %>%
  summarize(avg_mu_hat = mean(mu_hat))%>%
  ggplot(aes(x = mu, y = avg_mu_hat)) + geom_line() +
  xlab("True Mean")+
  ylab("Average Estimate of Mu Hat")

plot_rejected = 
  t_test_stats %>%
  mutate(
  reject = (p.value < 0.05)) %>%
  group_by(mu) %>%
  filter(reject == TRUE)%>%
  summarize(
    avg_reject = mean(mu_hat))%>%
  ggplot(aes(x = mu, y = avg_reject)) + geom_line() +
  xlab("True Mean")+
  ylab("Average Estimate of Mu Hat Where Null is Rejected")

plot_mu_hat + plot_rejected
  
```

**Is the sample average of μ hat across tests for which the null is rejected approximately equal to the true value of μ? Why or why not?**

ANSWER: Yes, the sample avergae is approximately equal to the true mean even when the null is rejected. This is because all the data was ultimately created from the same distribution with the same mean. The range of values of the estimate of mu hat is much larger when the null is rejected, but the mean value will still be the same. 