---
title: "Homework 5"
author: "Shritama Ray"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output: github_document

---

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)
library(broom)
library(p8105.datasets)
library(viridis)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

 

### Problem 2: Washington Post Homicides Data

**Read in & tidy the data:**
```{r read and tidy data}
#Read in data from github
homicides = read.csv('https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv') %>%
  janitor::clean_names()%>%
  mutate(
    city_state = str_c(city, ", ", state)
  )
```
The raw data set has *`r nrow(homicides)`* observations of *`r ncol(homicides)`* variables. It contains data on homicides, with variables such as  reported date,  victim's name,  race/age/sex,  disposition, city and state.

**Summary of Unsolved Homicides:**
```{r city summary}
homicides %>%
  filter(disposition == "Closed without arrest" | disposition == "Open/No arrest")%>%
  group_by(city_state, disposition)%>%
  count() %>%
  knitr::kable()
```

**Baltimore, MD Proportions**



### Problem 3: Power in a 1-sample t-test

n = 30
sigma = 5
mu = 0

**Generate 5000 datasets from the model: x ~ Normal[mu, sigma]:**
```{r sim mean sd}

#create function
sim_mean_pval = function(n, mu, sigma) {
  sim_data = tibble(
    x = rnorm(n, mean = mu, sd = sigma),
  )
  
  sim_data %>% 
    summarize(
      mu_hat = mean(x),
      t.test(x, mu = mu, alternative = "less", conf.level = 0.95) %>%
        broom::tidy() %>%
        select(p.value)
    )
}

#run function 5000 times and save results to a list
output = vector("list", 100)

for (i in 1:100) {
  output[[i]] = sim_mean_pval(n = 30, mu = 0, sigma = 5)
}

sim_results = bind_rows(output)
  
```

**Repeat the above for mu = 1, 2, 3, 4, 5, 6**
```{r varying mu}

#create a dataframe with all the outputs
t_test_stats =
tibble(
  mu = c(1:6)
) %>%
  mutate(
    stats = map(.x = mu, ~rerun(100, sim_mean_pval(n = 30, mu = .x, sigma = 5)))
  ) %>%
unnest(stats) %>%
unnest(stats)

```

**Plot proportion of times null was rejected vs. mu**
```{r plot power}
t_test_stats %>%
  mutate(
  reject = (p.value < 0.05)) %>%
  group_by(mu) %>%
    summarize(
    power = sum(reject)/length(p.value)) %>%
  ggplot(aes(x = mu, y = power)) + geom_line() +
  xlab("True Mean")+
  ylab("Test Power")
```

**Plot the average estimate of μ-hat vs. the true value of μ. Make a second plot of the average estimate of μ-hat only in samples for which the null was rejected vs. the true value of μ.**
```{r plot mu hat}
plot_mu_hat = 
  t_test_stats %>%
  group_by(mu) %>%
  summarize(avg_mu_hat = mean(mu_hat))%>%
  ggplot(aes(x = mu, y = avg_mu_hat)) + geom_line() +
  xlab("True Mean")+
  ylab("Average Estimate of Mu Hat")

plot_rejected = 
  t_test_stats %>%
  mutate(
  reject = (p.value < 0.05)) %>%
  group_by(mu) %>%
  filter(reject == TRUE)%>%
  summarize(
    avg_reject = mean(mu_hat))%>%
  ggplot(aes(x = mu, y = avg_reject)) + geom_line() +
  xlab("True Mean")+
  ylab("Average Estimate of Mu Hat Where Null is Rejected")

plot_mu_hat + plot_rejected
  
```

**Is the sample average of μ hat across tests for which the null is rejected approximately equal to the true value of μ? Why or why not?**

ANSWER: Yes, the sample avergae is approximately equal to the true mean even when the null is rejected. This is because all the data was ultimately created from the same distribution with the same mean. The range of values of the estimate of mu hat is much larger when the null is rejected, but the mean value will still be the same. 